---
title: "Qualitative Activity Recognition"
author: "ncapps"
date: "April 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Executive Summary
The aim of this project is to assess the quality of execution of weight lifting exercises using a model-based approach. Accelerometer data on the belt, forearm, arm, and dumbell from 6 participants was used to develop a model to predict the manner in which a participant completed a barbell lift.

### Loading and cleaing data
Data not measured by an accelerometer and variables where the majority of rows contained missing values were removed from the training dataset and were not used as predictors.
```{r}
library(AppliedPredictiveModeling);library(caret)
trainfilename <- "pml-training.csv"; testfilename <- "pml-testing.csv"
## Loaded 1622 rows and 160 columns
df <- read.csv(trainfilename,
                     header=TRUE,sep=",",na.strings=c("NA","#DIV/0!"))
## Removed 100 predictors where 97% of all values are NA : 60 columns remaining
df <- Filter(function(x){
        sum(is.na(x))/nrow(df) < 0.97}, df)
df <- df[,8:60] ## Remove non-accelerometer data
inTrain <- createDataPartition(df$classe, p = 3/4)[[1]]
training <- df[inTrain,]
testing <- df[-inTrain,]
set.seed(333)
```

### Model selection and cross Validation
The Random Forest algorithm was selected because there were a large number of predictors where the interactions between variables is unknown. The model was trained using 10-fold cross validation with 3 repeats. The out-of-bag error rate (OOB) was estimated to be 3.16%.

Training data was split such that 75% was used for training the model and 25% was used to evaluate model performance further. The random forest model had an accuracy of 98.71% on the test dataset.

```{r}
tc <- trainControl(method="repeatedcv", number=10, repeats=3)
rf.fit <- train(classe ~., method="rf", trControl=tc, ntree=10, data=training)
rf.fit$finalModel

rf.pred.test <- predict(rf.fit, testing)
confusionMatrix(rf.pred.test, testing$classe)$overall[1] # Accuracy = 0.9871533
```

### Predicting the Validation Test Set
```{r}
library(dplyr)
vdf <- read.csv(testfilename,header=TRUE,sep=",",na.strings=c("NA","#DIV/0!"))
vdf <- select(vdf, one_of(names(training[,-53])))
vdf <- rbind(training[1,-53], vdf);vdf <- vdf[-1,]
rf.pred.val <- predict(rf.fit, newdata=vdf)
rf.pred.val[1:20]
```

### References
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz4ejKqMRpu
